---
title: 'Racism - county - homeownership inequity'
author: "David Van Riper"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
require(tinytex)
require(gt)
require(readr)

topic <- "homeownership-inequity"
```

## Contents
* Description
* Source of Measure
* Input Data
* Data Tables
* Geographic Detail
* Temporal Detail
* Formula
* Interpretation
* Data Summary
* Citation and Use

## Description
Homeownership inequity is the ratio between the proportion of householders identifying as White alone, not Hispanic or Latino, who own (as opposed to renting) their home and the proportion of householders identifying as a different race/ethnic group who own their home. We have computed the homeownership inequity measure for the following groups:

* White alone, not Hispanic or Latino:Black alone
* White alone, not Hispanic or Latino:Asian alone
* White alone, not Hispanic or Latino:Hispanic or Latino

## Source of Measure
Chantarat, Tongtan, David C. Van Riper, and Rachel R. Hardeman. "Multidimensional structural racism predicts birth outcomes for Black and White Minnesotans." _Health Services Research_ 57, no. 3: 448-457. doi:10.1111/1475-6773.13976 

## Input Data
Steven Manson, Jonathan Schroeder, David Van Riper, Katherine Knowles, Tracy Kugler, Finn Roberts, and Steven Ruggles. IPUMS National Historical Geographic Information System: Version 18.0 [dataset]. Minneapolis, MN: IPUMS. 2023. http://doi.org/10.18128/D050.V18.0

We use data from the U.S. Census Bureau's American Community Survey 5-year summary files. 

## Data Tables
The following IPUMS NHGIS tables should be requested when creating the input dataset:

```{r, echo=FALSE, message=FALSE}
# Load the table code and label information
table_details <- read_csv(here::here("measures", topic, "data", "input", "table_code_label.csv"))

# Create a gt() table from table_details
table_details |> 
  gt() |> 
  cols_label(table_id = "Table ID",
             table_name = "Table name") |> 
  cols_width(
    table_name ~ px(400)
  )
  
```

## Geographic Detail

### Geographic Extent
Data are available for the United States.

### Geographic Units
Data are available for counties or county equivalents in the United States.

## Temporal Detail

### Temporal Extent
The American Community Survey (ACS) 5-year summary data are released every year, and each 5-year summary file contains survey responses pooled over a 5-year time span. For example, the 2015-2019 ACS 5-year data include survey responses collected from 2015-2019. The 2016-2020 ACS 5-year data include survey responses collected from 2016-2020. 

In order to generate homeownership inequity measures for all counties in the United States, we must use the 5-year summary data. Other ACS data products, including the 1-year summary data, only include estimates for geographic units whose population exceeds a certain threshold. 

## Formula 
We compute the educational inequity ratio using the following formula:

$$
homeown_{inequity_{ijk}} = \frac{prop_{homeown_{ij}}}{prop_{homeown_{ik}}} 
$$

where $homeown_{inequity_{ijk}}$ is the homeownership inequity ratio in county $i$ between groups $j$ and $k$, $prop_{homeown_{ij}}$ is the proportion of householders in group $j$ who own their home in county $i$, and $prop_{homeown_{ik}}$ is the proportion of householders in group $k$ who own their home in county $i$.

The group-specific homeownership proportions are computed using the following formula:

$$
prop_{homeown_{ij}} = \frac{N_{homeown_ij}}{N_{ij}}
$$

where $prop_{homeown_{ij}}$ is the proportion of householders in group $j$ who own their home in county $i$, $N_{homeown_ij}$ is the count of householders in group $j$ who own their house in county $i$, and $N_{ij}$ is the count of all householders in group $j$ in county $i$. 

## Interpretation
This measure compares the proportion of householders in a particular race/ethnicity category who own their house with the proportion of householders in a different race/ethnicity category who own their house. If the proportion for the numerator race/ethnicity category is greater than the proportion for the denominator race/ethnicity category, the ratio will be greater than 1.0. If the proportion for the numerator race/ethnicity category is less than the proportion for the denominator race/ethnicity category, the ratio will be less than 1.0.

For the measures in the data file, we use the White alone, not Hispanic or Latino category as the numerator for all the homeownership inequity ratios, and the denominator is one of the following three categories:

* Black or African American alone
* Asian alone
* Hispanic or Latino

Since the numerator and denominator are both percentages, the overall count of householders for a specific race/ethnicity category has less impact on the resulting ratio. 

## Missing data
We are unable to compute homeownership inequity ratios for some county - race/ethnicity combinations. If both the numerator and denominator proportions are zero, the homeownership inequity ratio is undefined. If the denominator proportion is zero, then the ratio is also undefined. When we observe these cases, we report the ratios as missing.  

## Data Summary
```{r, echo=FALSE, message=FALSE}
# Load the variable name, description and data type information
variable_details <- read_csv(here::here("measures", topic, "data", "input", "variable_descriptions.csv"))

# Create a gt() table from variable_details
variable_details |> 
  gt() |> 
  cols_label(variable_name = "Variable name",
             variable_description = "Variable description",
             data_type = "Data type") |> 
  cols_width(
    variable_description ~ px(300)
  )
```

## Citation and Use
All persons are granted a limited license to use this documentation and the
accompanying data, subject to the following conditions:

* REDISTRIBUTION: You will not redistribute the data without permission.

  You may publish a subset of the data to meet journal requirements for accessing 
  data related to a particular publication. Contact us for permission for any other 
  redistribution; we will consider requests for free and commercial redistribution.

* CITATION: You will cite CDOH appropriately.

Publications and research reports employing this measure should include the following citation:

* Rachel Hardeman, Claire Kamp Dush, Wendy Manning, and David Van Riper. Racism - county - homeownership inequity. IPUMS Contextual Determinants of Health. Minneapolis, MN: IPUMS. 2023. https://doi.org/10.18128/M130-011.2023-04

For policy briefs, online resources, or articles in the popular press, we recommend that you cite the use of CDOH data as follows:

* IPUMS CDOH, University of Minnesota, cdoh.ipums.org.

These terms of use are a legally binding agreement. You can use the data only in 
accordance with these terms, and any other use is a violation of the agreement. 
Violations may result in revocation of the agreement and prohibition from using other 
IPUMS data. If IPUMS or our partners are harmed from your violation, you are 
responsible for all damages, including reasonable attorney's fees and expenses.
        
In addition, we request that users send us a copy of any publications, research
reports, or educational material making use of the data or documentation.

Send electronic material to: ipums@umn.edu

Printed matter should be sent to:

    IPUMS CDOH
    Institute for Social Research and Data Innovation
    University of Minnesota
    50 Willey Hall
    225 19th Ave S
    Minneapolis, MN 55455
