---
title: "Sexism - state - paid family & medical leave"
author: "David Van Riper"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
require(tinytex)
require(gt)
require(readr)

topic <- "sexism-state-paid-family-medical-leave"
```

## Contents
* Description
* Source of Measure
* Input Data
* Geographic Detail
* Temporal Detail
* Formula
* Interpretation
* Data Summary
* Citation and Use

## Description
The state-level paid family & medical leave measure denotes whether a state has a law that guarantees paid family & medical leave for employees. 

## Source of Measure
Rapp, Kristen Schorpp, Vanessa V. Volpe, and Hannah Neukrug. "State-level sexism and women’s health care access in the United States: Differences by race/ethnicity, 2014–2019." _American Journal of Public Health_ 111, no. 10 (2021): 1796-1805.

## Input Data
National Partnership for Women & Families. State Paid Family & Medical Leave Insurance Laws. Washington, DC: 2023. https://www.nationalpartnership.org/our-work/resources/economic-justice/paid-leave/state-paid-family-leave-laws.pdf  

## Geographic Detail

### Geographic Extent
Data are available for the United States.

### Geographic Units
Data are available for each US state plus the District of Columbia.

## Temporal Detail

### Temporal Extent
The report lists the years in which paid family & medical leave insurance laws were enacted and went into effect. We determined whether a state had a law in effect from 2004 through 2023. The dataset starts in 2004 because that was the first year with an state-level family & medical leave in effect (California).  

### Temporal Frequency
Our measure is available on a yearly basis for 2004-2023. 

## Formula 
For a state $i$ in year $j$, we assign a value of 1 if there is a paid family & medical leave insurance law in effect. We assign a value of 0 if there is no paid family & medical insurance law in effect. 

## Interpretation
The paid family & medical leave variable indicates the presence of a law mandating such leave. The original data source provides details about the _enacted_ and _effective_ dates of the laws. We use the effective date of the law for this variable. The _effective date_ measures when individuals had access to the paid leave, as opposed to the _enacted date_ which measures when a state's government had the political will necessary to enact the policy. 

The original data source lists updates to paid family & medical leave policies, and these updates are typically expansions of existing policies. We do not capture the years of the updates since the original policies were already in effect.

The original data source provides a considerable amount of detail about each state's paid family & medical leave policies, including possible reasons for paid leave, definition of a family member, maximum length of paid leave, employee eligibility requirements, size of employer covered, and the benefit amount. Scientists may use these additional details to make finer grained distinctions among states with paid family & medical leave policies.    

Our coding system differs from Rapp et al. We use a value of 1 to indicate the existence of an effective paid family & medical leave law, while Rapp et al. used a value of 0 to indicate the existence of such laws. 

## Data Summary
```{r, echo=FALSE, message=FALSE}
# Load the variable name, description and data type information
variable_details <- read_csv(here::here("measures", topic, "data", "input", "variable_descriptions.csv"))

# Create a gt() table from variable_details
variable_details |> 
  gt() |> 
  cols_label(variable_name = "Variable name",
             variable_description = "Variable description",
             data_type = "Data type") |> 
  cols_width(
    variable_description ~ px(250)
  )
```
## Citation and Use
All persons are granted a limited license to use this documentation and the
accompanying data, subject to the following conditions:
  
  * REDISTRIBUTION: You will not redistribute the data without permission.

You may publish a subset of the data to meet journal requirements for accessing 
data related to a particular publication. Contact us for permission for any other 
redistribution; we will consider requests for free and commercial redistribution.

* CITATION: You will cite CDOH appropriately.

Publications and research reports employing this measure should include the following citation:
  
  * Claire Kamp Dush, Wendy Manning, and David Van Riper. Sexism - state - paid family & medical leave. IPUMS Contextual Determinants of Health. Minneapolis, MN: IPUMS. 2024. https://doi.org/10.18128/M130-004.2024-02

For policy briefs, online resources, or articles in the popular press, we recommend that you cite the use of CDOH data as follows:
  
  * IPUMS CDOH, University of Minnesota, cdoh.ipums.org.

These terms of use are a legally binding agreement. You can use the data only in 
accordance with these terms, and any other use is a violation of the agreement. 
Violations may result in revocation of the agreement and prohibition from using other 
IPUMS data. If IPUMS or our partners are harmed from your violation, you are 
responsible for all damages, including reasonable attorney's fees and expenses.
        
In addition, we request that users send us a copy of any publications, research
reports, or educational material making use of the data or documentation.

Send electronic material to: ipums@umn.edu

Printed matter should be sent to:

    IPUMS CDOH
    Institute for Social Research and Data Innovation
    University of Minnesota
    50 Willey Hall
    225 19th Ave S
    Minneapolis, MN 55455
