---
title: 'Politics - state - presidential election results'
author: "David Van Riper"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
require(tinytex)
require(gt)
require(readr)

topic <- "politics-state-presidential-election-results"
```

## Contents
* Description
* Input Data
* Geographic Detail
* Temporal Detail
* Formula
* Interpretation
* Data Summary
* Citation and Use

## Description
The state presidential election results measure provides the proportion of a state's votes cast for the Democratic party or the Republican party in presidential elections from 1976-2020.  

## Input Data
MIT Election Data and Science Lab, 2017, "U.S. President 1976–2020", https://doi.org/10.7910/DVN/42MVDX, Harvard Dataverse, V6, UNF:6:4KoNz9KgTkXy0ZBxJ9ZkOw== [fileUNF]

## Geographic Detail

### Geographic Extent
Data are available for the United States.

### Geographic Units
Data are available for each US state and the District of Columbia.

## Temporal Detail

### Temporal Extent
The "U.S. President 1976-2020" dataset provides state-level election returns from 1976-2020. 

### Temporal Frequency
Data are available on an every four-year basis from 1976-2020. 

## Formula 
We compute the proportion of votes received by the Democratic or Republican party using the following formula:

$$
Prop_{ij} = \frac{Votes_{ij}}{Votes_i} 
$$
where $Prop_{ij}$ is the proportion of state $i$'s votes cast for party $j$, $Votes_{ij}$ are the number of votes cast in state $i$ for party $j$, and $Votes_i$ is the total number of votes cast in state $i$.  

## Interpretation
This measure represents the state-level proportion of votes cast for the Democratic or Republican party in each presidential election from 1976-2020.

## Data Summary
```{r, echo=FALSE, message=FALSE}
# Load the variable name, description and data type information
variable_details <- read_csv(here::here("measures", topic, "data", "input", "variable_descriptions.csv"))

# Create a gt() table from variable_details
variable_details |> 
  gt() |> 
  cols_label(variable_name = "Variable name",
             variable_description = "Variable description",
             data_type = "Data type") |> 
  cols_width(
    variable_description ~ px(300)
  )
```

## Citation and Use
All persons are granted a limited license to use this documentation and the
accompanying data, subject to the following conditions:

* REDISTRIBUTION: You will not redistribute the data without permission.

  You may publish a subset of the data to meet journal requirements for accessing 
  data related to a particular publication. Contact us for permission for any other 
  redistribution; we will consider requests for free and commercial redistribution.

* CITATION: You will cite CDOH appropriately.

Publications and research reports employing this measure should include the following citation:

* Claire Kamp Dush, Wendy Manning, and David Van Riper. Politics - state - presidential election results. IPUMS Contextual Determinants of Health. Minneapolis, MN: IPUMS. 2024. https://doi.org/10.18128/M130-014.2024-02

For policy briefs, online resources, or articles in the popular press, we recommend that you cite the use of CDOH data as follows:

* IPUMS CDOH, University of Minnesota, cdoh.ipums.org.

These terms of use are a legally binding agreement. You can use the data only in 
accordance with these terms, and any other use is a violation of the agreement. 
Violations may result in revocation of the agreement and prohibition from using other 
IPUMS data. If IPUMS or our partners are harmed from your violation, you are 
responsible for all damages, including reasonable attorney's fees and expenses.
        
In addition, we request that users send us a copy of any publications, research
reports, or educational material making use of the data or documentation.

Send electronic material to: ipums@umn.edu

Printed matter should be sent to:

    IPUMS CDOH
    Institute for Social Research and Data Innovation
    University of Minnesota
    50 Willey Hall
    225 19th Ave S
    Minneapolis, MN 55455
