---
title: 'Sexism - state - proportion of state legislators identifying as female'
author: "David Van Riper"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
require(tinytex)
require(gt)
require(readr)

topic <- "sexism-state-proportion-state-legislators-identifying-as-female"
```

## Contents
* Description
* Source of Measure
* Input Data
* Geographic Detail
* Temporal Detail
* Formula
* Interpretation
* Notes
* Data Summary
* Citation and Use

## Description
The measure captures the proportion of state legislators who identify as female. We compute the proportion for the state legislature as a whole and for the state house and senate legislative chambers.  

## Source of Measure
Rapp, Kristen Schorpp, Vanessa V. Volpe, and Hannah Neukrug. "State-level sexism and women’s health care access in the United States: Differences by race/ethnicity, 2014–2019." _American Journal of Public Health_ 111, no. 10 (2021): 1796-1805.

## Input Data
The Center for American Women and Politics (CAWP) at Rutgers University-New Brunswick collects and publishes annual data from 2015-2022 on the number of female-identifying state legislators by state. The specific datasets we used to compute the percentages are: 

* Center for American Women and Politics (CAWP). 2024. “Women in State Legislatures 2023.” New Brunswick, NJ: Center for American Women and Politics, Eagleton Institute of Politics, Rutgers University-New Brunswick. https://cawp.rutgers.edu/facts/levels-office/state-legislature/women-state-legislatures-2024. (Accessed January 5, 2024) 

* Center for American Women and Politics (CAWP). 2022. “Women in State Legislatures 2022.” New Brunswick, NJ: Center for American Women and Politics, Eagleton Institute of Politics, Rutgers University-New Brunswick. https://cawp.rutgers.edu/facts/levels-office/state-legislature/women-state-legislatures-2022. (Accessed August 31, 2022) 

* Center for American Women and Politics (CAWP). 2021. “Women in State Legislatures 2021.” New Brunswick, NJ: Center for American Women and Politics, Eagleton Institute of Politics, Rutgers University-New Brunswick. https://cawp.rutgers.edu/facts/levels-office/state-legislature/women-state-legislatures-2021. (Accessed August 31, 2022) 

* Center for American Women and Politics (CAWP). 2020. “Women in State Legislatures 2020.” New Brunswick, NJ: Center for American Women and Politics, Eagleton Institute of Politics, Rutgers University-New Brunswick. https://cawp.rutgers.edu/facts/levels-office/state-legislature/women-state-legislatures-2020. (Accessed August 31, 2022) 

* Center for American Women and Politics (CAWP). 2019. “Women in State Legislatures 2019.” New Brunswick, NJ: Center for American Women and Politics, Eagleton Institute of Politics, Rutgers University-New Brunswick. https://cawp.rutgers.edu/facts/levels-office/state-legislature/women-state-legislatures-2019. (Accessed August 31, 2022) 

* Center for American Women and Politics (CAWP). 2018. “Women in State Legislatures 2018.” New Brunswick, NJ: Center for American Women and Politics, Eagleton Institute of Politics, Rutgers University-New Brunswick. https://cawp.rutgers.edu/facts/levels-office/state-legislature/women-state-legislatures-2018. (Accessed August 31, 2022) 

* Center for American Women and Politics (CAWP). 2017. “Women in State Legislatures 2017.” New Brunswick, NJ: Center for American Women and Politics, Eagleton Institute of Politics, Rutgers University-New Brunswick. https://cawp.rutgers.edu/facts/levels-office/state-legislature/women-state-legislatures-2017. (Accessed August 31, 2022) 

* Center for American Women and Politics (CAWP). 2016. “Women in State Legislatures 2016.” New Brunswick, NJ: Center for American Women and Politics, Eagleton Institute of Politics, Rutgers University-New Brunswick. https://cawp.rutgers.edu/facts/levels-office/state-legislature/women-state-legislatures-2016. (Accessed August 31, 2022) 

* Center for American Women and Politics (CAWP). 2015. “Women in State Legislatures 2015.” New Brunswick, NJ: Center for American Women and Politics, Eagleton Institute of Politics, Rutgers University-New Brunswick. https://cawp.rutgers.edu/facts/levels-office/state-legislature/women-state-legislatures-2015. (Accessed August 31, 2022) 

## Geographic Detail

### Geographic Extent
Data are available for the United States.

### Geographic Units
Data are available for each US state and the District of Columbia. 

## Temporal Detail

### Temporal Extent
Data are available from 2015-2023. 

### Temporal Frequency
Data are available on an annual basis. 

## Formula 
We compute the proportion of state legislators identifying as female using the following formula:

$$
Proportion_{female_{ij}}  = \frac{Legislators_{female_{ij}}}{Legislators_{total_{ij}}}
$$

where the $Proportion_{female_{ij}}$  is the percentage of state $i$ legislators identifying as female in year $j$. The numerator is number of state $i$ legislators identifying as female in year $j$. The denominator is the number of legislators in state $i$ in year $j$.

We compute house or senate chamber-specific percentages using the following formula:


$$
Proportion_{female_{ijk}}  = \frac{Legislators_{female_{ijk}}}{Legislators_{total_{ijk}}} 
$$
where $Proportion_{female_{ijk}}$  is the percentage of state $i$ legislators identifying as female in year $j$ and legislative chamber $k$. The numerator is the number of state $i$ legislators identifying as female in year $j$ and legislative chamber $k$. The denominator is the number of legislators in state $i$ in year $j$ and legislative chamber $k$.

## Interpretation
This measure represents the proportion of a state's legislators who identify as female in a particular year. 

## Notes
Nebraska has a unicameral state legislature; therefore, it has proportions for _prop_female_legis_ and _prop_female_senate_. We set its _prop_female_house_ value to _0_.

## Data Summary
```{r, echo=FALSE, message=FALSE}
# Load the variable name, description and data type information
variable_details <- read_csv(here::here("measures", topic, "data", "input", "variable_descriptions.csv"))

# Create a gt() table from variable_details
variable_details |> 
  gt() |> 
  cols_label(variable_name = "Variable name",
             variable_description = "Variable description",
             data_type = "Data type") |> 
  cols_width(
    variable_description ~ px(250)
  )
```

## Citation and Use
All persons are granted a limited license to use this documentation and the
accompanying data, subject to the following conditions:

* REDISTRIBUTION: You will not redistribute the data without permission.

  You may publish a subset of the data to meet journal requirements for accessing 
  data related to a particular publication. Contact us for permission for any other 
  redistribution; we will consider requests for free and commercial redistribution.

* CITATION: You will cite CDOH appropriately.

Publications and research reports employing this measure should include the following citation:

* Claire Kamp Dush, Wendy Manning, and David Van Riper. Sexism - state - proportion of state legislators identifying as female. IPUMS Contextual Determinants of Health. Minneapolis, MN: IPUMS. 2023. https://doi.org/10.18128/M130-006.2023-04 

For policy briefs, online resources, or articles in the popular press, we recommend that you cite the use of CDOH data as follows:

* IPUMS CDOH, University of Minnesota, cdoh.ipums.org.

These terms of use are a legally binding agreement. You can use the data only in 
accordance with these terms, and any other use is a violation of the agreement. 
Violations may result in revocation of the agreement and prohibition from using other 
IPUMS data. If IPUMS or our partners are harmed from your violation, you are 
responsible for all damages, including reasonable attorney's fees and expenses.
        
In addition, we request that users send us a copy of any publications, research
reports, or educational material making use of the data or documentation.

Send electronic material to: ipums@umn.edu

Printed matter should be sent to:

    IPUMS CDOH
    Institute for Social Research and Data Innovation
    University of Minnesota
    50 Willey Hall
    225 19th Ave S
    Minneapolis, MN 55455
