---
title: 'Racism - county - income inequity'
author: "David Van Riper"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
require(tinytex)
require(gt)
require(readr)

topic <- "racism-county-income-inequity"
```

## Contents
* Description
* Source of Measure
* Input Data
* Data Tables
* Geographic Detail
* Temporal Detail
* Formula
* Interpretation
* Data Summary
* Citation and Use

## Description
Income inequity is measured using the index of concentration at the extremes (ICE). ICE, developed by Douglas Massey in 2001, is a measure of social polarization within a particular geographic unit. It shows whether people or households in a geographic unit are concentrated in privileged or deprived extremes. Our privileged group is the number of households with a householder identifying as White alone, not Hispanic or Latino, with an income equal to or greater than \$100,000. Our deprived group is the number of households with a householder identifying as a different race/ethnic group (e.g., Black alone, Asian alone, Hispanic or Latino), with an income equal to or less than \$25,000. 

We have computed the income inequity measure for the following groups:

* White alone, not Hispanic or Latino:Black alone
* White alone, not Hispanic or Latino:Asian alone
* White alone, not Hispanic or Latino:Hispanic or Latino

## Source of Measure
Chantarat, Tongtan, David C. Van Riper, and Rachel R. Hardeman. "Multidimensional structural racism predicts birth outcomes for Black and White Minnesotans." _Health Services Research_ 57, no. 3: 448-457. doi:10.1111/1475-6773.13976 

Massey, Douglas. "The Prodigal Paradigm Returns: Ecology Comes Back to Sociology." In _Does It Take a Village? Community Effects on Children, Adolescents, and Families_. Alan Booth and Ann C. Crouter (eds). Mahwewh, NJ: Lawrence Erlbaum Associates. 2001. 

## Input Data
Steven Manson, Jonathan Schroeder, David Van Riper, Katherine Knowles, Tracy Kugler, Finn Roberts, and Steven Ruggles. IPUMS National Historical Geographic Information System: Version 18.0 [dataset]. Minneapolis, MN: IPUMS. 2023. http://doi.org/10.18128/D050.V18.0

We use data from the U.S. Census Bureau's American Community Survey 5-year summary files. 

## Data Tables
The following IPUMS NHGIS tables should be requested when creating the input data:

```{r, echo=FALSE, message=FALSE}
# Load the table code and label information
table_details <- read_csv(here::here("measures", topic, "data", "input", "table_code_label.csv"))

# Create a gt() table from table_details
table_details |> 
  gt() |> 
  cols_label(table_id = "Table ID",
             table_name = "Table name") |> 
  cols_width(
    table_name ~ px(400)
  )
  
```

## Geographic Detail

### Geographic Extent
Data are available for the United States.

### Geographic Units
Data are available for counties or county equivalents in the United States.

## Temporal Detail

### Temporal Extent
The American Community Survey (ACS) 5-year summary data are released every year, and each 5-year summary file contains survey responses pooled over a 5-year time span. For example, the 2015-2019 ACS 5-year data include survey responses collected from 2015-2019. The 2016-2020 ACS 5-year data include survey responses collected from 2016-2020. 

In order to generate income inequity measures for all counties in the United States, we must use the 5-year summary data. Other ACS data products, including the 1-year summary data, only include estimates for geographic units whose population exceeds a certain threshold. 

## Formula 
We compute the income inequity measure using the following formula:

$$
ICE_{i} = \frac{A_{ij} - P_{ik}}{T_i}
$$

where $ICE_{i}$ is the income inequity value in county $i$, $A_{ij}$ is the count of households with a householder in group $j$ whose income is greater than or equal to \$100,000 in county $i$, $P_{ik}$ is the count of households with a householder in group $k$ whose income is less than or equal to \$25,000 in county $i$, and $T_i$ is the number of households in county $i$.

For our income inequity measures, group $j$ is householders identifying as White alone, not Hispanic or Latino, and group $k$ may be householders identifying as Black alone, Asian alone, or Hispanic or Latino. 

## Interpretation
This measure captures the relative concentration of groups in privilege or deprivation. Values for ICE range from -1 to 1, and the magnitudes of $A_{ij}$, $P_{ik}$, $T_i$ all influence the value of ICE. A few examples can help you interpret values of ICE you observe in the data file.

We observe a value of -1.0 if there are zero households in group $A_{ij}$ in a county and all other households in the county are in group $P_{ik}$. If all households in a county are in group $P_{ik}$, then $P_{ik}$ and $T_i$ are equal. If there are 500 households in a county ($T_i$) and all are in $P_{ik}$, then we will observe:

$$
-1.0 = \frac{0 - 500}{500}
$$

We observe a value of 1 if all households in a county are in group $A_{ij}$ in a county and zero households in the county are in group $P_{ik}$. If all households in a county are in group $A_{ij}$, then $A_{ij}$ and $T_i$ are equal. If there are 500 households in a county ($T_i$) and all are in $A_{ij}$, then we will observe:

$$
1.0 = \frac{500 - 0}{500}
$$

We observe a value of 0.0 if there are the same number of households in $A_{ij}$ and $P_{ik}$ in a county and there are no other households in the county. Thus, the sum of $A_{ij}$ and $P_{ik}$ equals $T_i$. If there are 500 households in a county ($T_i$) and they are evenly split between $A_{ij}$ and $P_{ik}$, then we will observe:

$$
0.0 = \frac{250 - 250}{500}
$$

We may also observe values close to 0.0 if both counts in the numerator are small relative to the count in the denominator. Similarly, we may observe values close to 1.0 if $A_{ij}$ and $T_i$ have similar magnitudes. If $P_{ik}$ and $T_i$ are similar, then we will observe values close to -1.0.

## Data Summary
```{r, echo=FALSE, message=FALSE}
# Load the variable name, description and data type information
variable_details <- read_csv(here::here("measures", topic, "data", "input", "variable_descriptions.csv"))

# Create a gt() table from variable_details
variable_details |> 
  gt() |> 
  cols_label(variable_name = "Variable name",
             variable_description = "Variable description",
             data_type = "Data type") |> 
  cols_width(
    variable_description ~ px(300)
  )
```

## Citation and Use
All persons are granted a limited license to use this documentation and the
accompanying data, subject to the following conditions:

* REDISTRIBUTION: You will not redistribute the data without permission.

  You may publish a subset of the data to meet journal requirements for accessing 
  data related to a particular publication. Contact us for permission for any other 
  redistribution; we will consider requests for free and commercial redistribution.

* CITATION: You will cite CDOH appropriately.

Publications and research reports employing this measure should include the following citation:

* Rachel Hardeman, Claire Kamp Dush, Wendy Manning, and David Van Riper. Racism - county - income inequity. IPUMS Contextual Determinants of Health. Minneapolis, MN: IPUMS. 2024. https://doi.org/10.18128/M130-012.2024-02

For policy briefs, online resources, or articles in the popular press, we recommend that you cite the use of CDOH data as follows:

* IPUMS CDOH, University of Minnesota, cdoh.ipums.org.

These terms of use are a legally binding agreement. You can use the data only in 
accordance with these terms, and any other use is a violation of the agreement. 
Violations may result in revocation of the agreement and prohibition from using other 
IPUMS data. If IPUMS or our partners are harmed from your violation, you are 
responsible for all damages, including reasonable attorney's fees and expenses.
        
In addition, we request that users send us a copy of any publications, research
reports, or educational material making use of the data or documentation.

Send electronic material to: ipums@umn.edu

Printed matter should be sent to:

    IPUMS CDOH
    Institute for Social Research and Data Innovation
    University of Minnesota
    50 Willey Hall
    225 19th Ave S
    Minneapolis, MN 55455
